---
title: "Linear models"
output: github_document
---

```{r setup, include = FALSE}
library(tidyverse)
library(tidyverse)
library(p8105.datasets)
library(broom)

set.seed(1)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


## Import data

```{r}
data("nyc_airbnb")

nyc_airbnb = 
  nyc_airbnb %>% 
  mutate(stars = review_scores_location / 2) %>% 
  rename(
    borough = neighbourhood_group,
    neighborhood = neighbourhood) %>% 
  filter(borough != "Staten Island") %>% 
  select(price, stars, borough, neighborhood, room_type)

fit = lm(price ~ stars + borough, data = nyc_airbnb)
```

## Model fitting

Let's look at the result....
Tidying output

```{r}
summary(fit)
summary(fit)$coef
coef(fit)
fitted.values(fit)
```

Let's look at the result better ....

```{r}
fit %>% 
  broom::glance()

fit %>% 
  broom::tidy()
```


```{r}
fit %>% 
  broom::tidy() %>% 
  select(term, estimate, p.value) %>% 
  mutate(term = str_replace(term, "^borough", "Borough: ")) %>% 
  knitr::kable(digits = 3)
```

## Be in control of factors

```{r}
nyc_airbnb = 
  nyc_airbnb %>% 
  mutate(
    borough = fct_infreq(borough), # change the frequence, and compare different groups/ bronx as reference changed to manhaton as reference.
    room_type = fct_infreq(room_type)) ####

fit = lm(price ~ stars + borough, data = nyc_airbnb)
```


## Diagnostics

```{r}
modelr::add_residuals(nyc_airbnb, fit)

modelr::add_predictions(nyc_airbnb, fit)
```

```{r}
nyc_airbnb %>% 
  modelr::add_residuals(fit) %>% 
  ggplot(aes(x = borough, y = resid)) +
  geom_violin() +
  ylim(-500, 1000) # add a limit to y values

nyc_airbnb %>% 
  modelr::add_residuals(fit) %>% 
  ggplot(aes(x = stars, y = resid, color = "magma")) + 
  geom_point() +
  facet_wrap(. ~ borough)
```

## Hypothesis testing

(doing t-test by default)

What about the significance of `borough`.

```{r}
fit_null = lm(price ~ stars + borough, data = nyc_airbnb)
fit_alt = lm(price ~ stars + borough + room_type, data = nyc_airbnb)

anova(fit_null, fit_alt) %>% 
  broom::tidy()
```

## Nesting data

This is pretty formal and also complex
```{r}
nyc_airbnb %>% 
  lm(price ~ stars * borough + room_type * borough, data = .) %>% 
  broom::tidy() %>% 
  knitr::kable(digits = 3)
```

This is more exploratory but maybe easier to understand.

```{r}
nest_lm_res =
  nyc_airbnb %>% 
  nest(data = -borough) %>% 
  mutate(
    models = map(data, ~lm(price ~ stars + room_type, data = .x)), # pull(models)
    results = map(models, broom::tidy)) %>% # get the tidy data
  select(-data, -models) %>% 
  unnest(results)

nest_lm_res %>% 
  select(borough, term, estimate) %>% 
  mutate(term = fct_inorder(term)) %>% 
  pivot_wider(
    names_from = term, values_from = estimate) %>% 
  knitr::kable(digits = 3) # increase in stars increase ..., going from entire to private and shared room save ... dolars.
```

Let's nest even more ...

```{r}
manhattan_airbnb =
  nyc_airbnb %>% 
  filter(borough == "Manhattan")

manhattan_nest_lm_res =
  manhattan_airbnb %>% 
  nest(data = -neighborhood) %>% 
  mutate(
    models = map(data, ~lm(price ~ stars + room_type, data = .x)),
    results = map(models, broom::tidy)) %>% 
  dplyr::select(-data, -models) %>% 
  unnest(results)

manhattan_nest_lm_res %>% 
  filter(str_detect(term, "room_type")) %>% 
  ggplot(aes(x = neighborhood, y = estimate)) + 
  geom_point() + 
  facet_wrap(~term) + 
  theme(axis.text.x = element_text(angle = 80, hjust = 1))
```


With this many factor levels, it really isn’t a good idea to fit models with main effects or interactions for each. Instead, you’d be best-off using a mixed model, with random intercepts and slopes for each neighborhood. Although it’s well beyond the scope of this class, code to fit a mixed model with neighborhood-level random intercepts and random slopes for room type is below. And, of course, we can tidy the results using a mixed-model spinoff of the broom package.  

```{r}
manhattan_airbnb %>% 
  lme4::lmer(price ~ stars + room_type + (1 + room_type | neighborhood), data = .) %>% 
  broom.mixed::tidy()
```


## Binary outcomes

baltimore_df = 
  read_csv("data/homicide-data.csv") %>% 
  filter(city == "Baltimore") %>% 
  mutate(
    resolved = as.numeric(disposition == "Closed by arrest"),
    victim_age = as.numeric(victim_age),
    victim_race = fct_relevel(victim_race, "White")) %>% 
  select(resolved, victim_age, victim_race, victim_sex)

fit_logistic = 
  baltimore_df %>% 
  glm(resolved ~ victim_age + victim_race + victim_sex, data = ., family = binomial())   

fit_logistic %>% 
  broom::tidy() %>% 
  mutate(OR = exp(estimate)) %>%
  select(term, log_OR = estimate, OR, p.value) %>% 
  knitr::kable(digits = 3)

baltimore_df %>% 
  modelr::add_predictions(fit_logistic) %>% 
  mutate(fitted_prob = boot::inv.logit(pred))